$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-05-15 paper/Referee-response.tex

The referee is correct to point out that there is a large theoretical
uncertainty on the Higgs mass and we did not mean to suggest
otherwise.  When we discuss the fine tuning dependence on the Higgs
mass we are trying to show the qualitative behavior of the tuning
increasing with the Higgs mass and how this changes depending on the
UV cutoff.  We agree that the theoretical error on the Higgs mass is
rather large and that this affects the impact the Higgs mass
measurement has on the minimum tuning that is consistent with
observation.  However such discussions raise difficult questions about
how large the theoretical error is and how to address that in the context
of fine tuning.  Therefore we leave such tricky questions to papers
which are focussed on the impact on the Higgs rather than the fine
tuning from the $Z^\prime$ mass which we are interested in here.

To address the referee's remarks we have adjusted the wording of the
first comment referred to and have added a footnote pointing out that
there will be a sizable uncertainty of the Higgs mass.

\item {\it ``looking at their table II, one sees that in certain parts of
the parameter space, $A_t$ can be much larger than the corresponding soft
masses. This implies the danger of color and charge breaking minima as is
known for long time. I guess that quite some points are excluded by this.
Moreover, it seems that they allow also $|\mu|$ to be below 100 GeV which
would conflict LEP2 data. Are such points really included in their analysis?''}

For parts of the parameter space scanned the referee is correct to point
out that there is a danger of color or charge breaking minima, and that
the range of values for $\mu$ scanned over includes values that would be in
conflict with experimental limits.  In our analysis we excluded points
that may lead to such minima by removing those which have tachyons in the
spectrum, and for the MSSM we have additionally checked that included points
satisfy known conditions for the dangerous minima to be avoided.  This however
was not commented on in our submission.  Regarding the LEP2 data and chargino
mass bounds, in Figures 2 and 4 the chargino mass limit was not imposed.
Instead the impact of the chargino bounds is explored in Figure 2.  As can
be inferred from this figure, and we confirmed that this is the case in
detail for all of the models considered, excluding low mass charginos does not
change the lower bound on fine tuning that is obtained.  Nevertheless the
referee is correct in pointing out that it is more consistent to present only
those points that are allowed by the experimental limits, as we have already
done for other limits on sparticle masses.

To address this point, we have updated Figures 2 and 4 to exclude those points
that would conflict with the LEP2 data.  As noted above, this does not alter
the results at all.  In addition, a short discussion has been added on page 17,
prior to Table II, in which the constraints that are applied to the analysis
are clarified.  In particular, we explicitly state that points violating the
LEP chargino limits are now excluded from the analysis.  This is accompanied
by a footnote pointing out that for the region of parameter space studied
whether or not this constraint is applied does not in fact impact the results.
We also explain that in order to exclude points that may have color and charge
breaking minima we only allow points that do not lead to tachyonic sparticle
masses and, in the MSSM, that satisfy the traditional conditions in the
literature for color or charge breaking minima to be avoided.

### 2015-05-16 paper/Referee-response.tex

The referee is correct to point out that there is a large theoretical
uncertainty on the Higgs mass and we did not mean to suggest
otherwise.  When we discuss the fine tuning dependence on the Higgs
mass we are trying to show the qualitative behavior of the tuning
increasing with the Higgs mass and how this changes depending on the
UV cutoff.  We agree that the theoretical error on the Higgs mass is
rather large and that this affects the impact the Higgs mass
measurement has on the minimum tuning that is consistent with
observation.  However such discussions raise difficult questions about
how large the theoretical error is and how to address that in the context
of fine tuning.  Therefore we leave such tricky questions to papers
which are focussed on the impact on the Higgs rather than the fine
tuning from the $Z^\prime$ mass which we are interested in here.

To address the referee's remarks we have adjusted the wording of the
first comment referred to and have added a footnote pointing out that
there will be a sizable uncertainty of the Higgs mass.

\item {\it ``looking at their table II, one sees that in certain parts of
the parameter space, $A_t$ can be much larger than the corresponding soft
masses. This implies the danger of color and charge breaking minima as is
known for long time. I guess that quite some points are excluded by this.
Moreover, it seems that they allow also $|\mu|$ to be below 100 GeV which
would conflict LEP2 data. Are such points really included in their analysis?''}

For parts of the parameter space scanned the referee is correct to point
out that there is a danger of color or charge breaking minima, and that
the range of values for $\mu$ scanned over includes values that would be in
conflict with experimental limits.  In our analysis we excluded points
that may lead to such minima by removing those which have tachyons in the
spectrum, and for the MSSM we have additionally checked that included points
satisfy known conditions for the dangerous minima to be avoided.  This however
was not commented on in our submission.  Regarding the LEP2 data and chargino
mass bounds, in Figures 2 and 4 the chargino mass limit was not imposed.  This
is because the LEP bounds are not severe enough to effect the
minimum obtainable fine tuning in the models we looked at.  On the other hand,
much higher limits set at the LHC can have a meaningful impact, as is
illustrated in  Figure 2.  Nevertheless the referee is correct in pointing out
that this should be addressed in the text.

To address these points, a short comment has been added on page 17,
prior to Table II, in which the theoretical constraints that are applied to
the analysis are clarified.  In particular, we explain that in order to
exclude points that may have color and charge breaking minima we only allow
points that do not lead to tachyonic sparticle masses and, in the MSSM, that
satisfy the traditional conditions in the literature for color or charge
breaking minima to be avoided.  In addition to this, we have added a remark
on page 18 to point out that imposing the LEP bound does not have an effect
on the results, but that LHC limits can have a significant impact.  We also
have added a footnote stating that it is for this reason the LEP bounds are
not imposed in Figures 2 and 4.

### 2015-05-16 paper/e6ssm-tuning-paper.tex
Modified comment on LEP bounds in paper and referee response

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  In order to avoid being in conflict with LEP
bounds on the chargino mass \cite{Kraan:2005vy} we exclude those points for
which the lightest chargino mass $m_{\tilde{\chi}^\pm_1} < 104$ GeV\footnote{In
fact this requirement has no impact on the bounds we find for the fine tuning
below. We found that, in the parameter space covered by our scans, including
points that violate the LEP bound does not lower the minimum fine tuning that
is obtained}.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
As a further requirement in the MSSM we exclude any points that do not satisfy
the well known CCB constraint \cite{Casas:1995pd},
\begin{equation} \label{eq:MSSMCCBconstraint}
A_t^2 < 3 \left ( \mu^2 + m_{H_u}^2 + m_{Q_3}^2 + m_{u_3}^2 \right ),
\end{equation}
or the unbounded from below (UFB) constraint,
\begin{equation} \label{eq:MSSMUFBconstraint}
m_{H_d}^2 + m_{H_u}^2 + 2 \mu^2 > 2 B \mu .
\end{equation}
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  However the chargino limits from the LHC
depend on whether there are light sleptons or sneutrinos and the mass
difference between the lightest chargino and lightest neutralino.
Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.

>>>>

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
As a further requirement in the MSSM we exclude any points that do not satisfy
the well known CCB constraint \cite{Casas:1995pd},
\begin{equation} \label{eq:MSSMCCBconstraint}
A_t^2 < 3 \left ( \mu^2 + m_{H_u}^2 + m_{Q_3}^2 + m_{u_3}^2 \right ),
\end{equation}
or the unbounded from below (UFB) constraint,
\begin{equation} \label{eq:MSSMUFBconstraint}
m_{H_d}^2 + m_{H_u}^2 + 2 \mu^2 > 2 B \mu .
\end{equation}
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  The LEP bound \cite{Kraan:2005vy} on
chargino masses, excluding $m_{\tilde{\chi}^\pm_1} \lesssim 104$ GeV, implies
that $|\mu|$ should only be greater than $\sim 100$ GeV, which is not
substantially larger than $M_Z$.  Consequently the bound from LEP is not
high enough to have an impact on the fine tuning obtained in the models and
parameter space regions that we have studied, as we have checked explicitly
\footnote{It is for this reason that we did not require the LEP bounds be
satisfied for the points shown in Figure \ref{Fig:e6ssmvsmssm}
and in Figure \ref{Fig:othere6modelsvsmssm} below.}.  Significantly larger
lower bounds on the $\mu$ parameter, and therefore on the fine tuning, may
arise from chargino limits coming from LHC searches.  However the chargino
limits from the LHC depend on whether there are light sleptons or sneutrinos
and the mass difference between the lightest chargino and lightest neutralino.
Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.
